The Astropy Project:

### 2017-09-28 main.tex
Update on Overleaf.

\section{Introduction} \label{sec:intro}
% first draft by Moritz (hamogu)
All astronomical research makes use of software in some way and
astronomy as a field has long supported the development of specialized
software tools for different astronomical tasks. Some of those
software packages are (or at least were) supported by large
institutions, e.g.\ IRAF \citep[developed at NAOA,][]{IRAF}, MIDAS
\citep[developed at ESO,][]{MIDAS}, or ds9 \citep[developed at
  SAO,][]{ds9}. These packages tend to address a wide range of users
and they typically provide some level of documentation and user
support. Other packages are developed foremost by individual research
groups and used mostly in work by members of this group. Those
packages can address more specialized needs, but are not necessarily
available to the community and even if they are, documentation can in
some cases be insufficient for new users and the quality of the
software, i.e. the robustness and the level of testing that has been
performed to ensure correctness, is not always clear. On the other
hand, there are also many widely used, well documented, high quality
software packages which are developed by individuals or small groups.

@@

At the same time, an ecosystem of tools, programms, and webservices
has become available that makes it easier to collaborate on code development,
continuously test the code, build and provide documentation, and
supply downloadable packages. The astropy project aims to develop and
provide high quality code and documentation according to the best
practices in software development. The project makes use of these
tools to do so without central institutional oversight.

The first public release of the astropy package is described in
\cite{2013A&A...558A..33A}. Since then, the astropy package has been
used in hundreds of projects, and the scope of the package as grown
considerably. At the same time, the community of astronomers
contributing to the project has grown tremendeously and an ecosystem
of packages supporting or affiliated with the astropy core has
developed. In this paper, we describe the current status of the
astropy community, the astropy core package, and the astropy
ecosystem.

We start by describing the way the astropy project works and is
organized in section~\ref{sec:concepts}.  The project develops a core
package called \texttt{astropy} (section~\ref{sec:core}) and several
separate packages that help setting up the infrastructure for testing,
documentation etc.\ (section~\ref{sec:infrastructure}). In
section~\ref{sec:learning} we give an overview of the wealth of
recources created to teach new users the use of the astropy packages
and how to contribute to the community. We end with a short vision for
the future of astropy in section~\ref{sec:future}.

@@

@@

\subsection{Statistics}
% Steve C.
% Jake V.:  Lomb-scargle, Bayesian blocks
% Larry:  sigma clipping, biweight stats
major additions to be discussed: lomb-scargle, sigma clipping, bayesian blocks, histograms, also:  robust (biweight) statistics

@@

\section{Learning Astropy}
\label{sec:learning}
% Kelle, Adrian

>>>>

\section{Introduction} \label{sec:intro}
% first draft by Moritz (hamogu)
All astronomical research makes use of software in some way and
astronomy as a field has long supported the development of specialized
software tools for different astronomical tasks. Some of those
software packages are (or at least were) supported by large
institutions, e.g.\ IRAF \citep[developed at NOAO,][]{IRAF}, MIDAS
\citep[developed at ESO,][]{MIDAS}, or ds9 \citep[developed at
% * <bsipocz@gmail.com> 2017-09-28T21:05:00.569Z:
% 
% > or ds9 
% probably starlink is a better example as being a large package of software similarly to IRAF and MIDAS rather than a single tool.
% 
% ^.
  SAO,][]{ds9}. These packages tend to address a wide range of users
and they typically provide some level of documentation and user
support. Other packages are developed foremost by individual research
groups and used mostly in work by members of this group. Those
packages can address more specialized needs, but are not necessarily
available to the community and even if they are, documentation can in
some cases be insufficient for new users and the quality of the
software, i.e. the robustness and the level of testing that has been
performed to ensure correctness, is not always clear. On the other
hand, there are also many widely used, well documented, high quality
software packages which are developed by individuals or small groups.

@@

At the same time, an ecosystem of tools, programs, and webservices
has become available that makes it easier to collaborate on code development,
continuously test the code, build and provide documentation, and
supply downloadable packages. The astropy project aims to develop and
provide high quality code and documentation according to the best
practices in software development. The project makes use of these
tools to do so without central institutional oversight.

The first public release of the astropy package is described in
\cite{2013A&A...558A..33A}. Since then, the astropy package has been
used in hundreds of projects, and the scope of the package as grown
considerably. At the same time, the community of astronomers
contributing to the project has grown tremendously and an ecosystem
of packages supporting or affiliated with the astropy core has
developed. In this paper, we describe the current status of the
astropy community, the astropy core package, and the astropy
ecosystem.

We start by describing the way the astropy project works and is
organized in section~\ref{sec:concepts}.  The project develops a core
package called \texttt{astropy} (section~\ref{sec:core}) and several
separate packages that help setting up the infrastructure for testing,
documentation etc.\ (section~\ref{sec:infrastructure}). In
section~\ref{sec:learning} we give an overview of the wealth of
resources created to teach new users the use of the astropy packages
and how to contribute to the community. We end with a short vision for
the future of astropy in section~\ref{sec:future}.

@@

\subsubsection{Unified file read/write interface}

Check when it was added (0.3?). Overview of the supported formats (ascii, fits, votable).

\subsubsection{Enhanced CSV format}

\subsubsection{ASCII, FITS, etc.}

One subsection per format (if enough new features) ?

\begin{itemize}
\item ascii : fast reader/writer (1.0), HTML tables
\item fits: lazy loading, command-line scripts (fitsheader, fitsinfo)
\end{itemize}

@@

\subsection{Statistics}
% Steve C.: overview, circular stats, 
% Jake V.:  Lomb-scargle, Bayesian blocks
% Larry:  sigma clipping, biweight stats
% Ze':  Ripley's K (spatial stats)
major additions to be discussed: lomb-scargle, sigma clipping, bayesian blocks, histograms, also:  robust (biweight) statistics

\subsection{Robust Statistical Estimators}
\subsection{Circular Statistics}
\subsection{Lomb-Scargle Periodograms}
\subsection{Ripley's K Function Estimators}
\subsection{Histogram Binning}

@@

\section{Learning Astropy}
\label{sec:learning}
% Kelle, Adrian
% - would astropy learn mature enough to described in the paper, or maybe it should go as a subsection for the next, future of astropy section?
